The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2014-02-13 paper/ms.tex
font size in histograms
rewriting the conclusions

We use the full width at half maximum (FWHM) to quantify the line
broadening. Figure \ref{fig:widthvsvelocity} shows how $FWHM$ increases
with rotational velocity both for the central an homogeneous source
distributions.  We parameterize the dependency of the linewidth as
\begin{equation}
 FWHM^2 = FWHM_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where $FWHM_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.9 \pm 3.17\times 10^{-17}$ and $\lambda_{\rm h}=
1.9\pm 3.07\times 10^{-10}$ respectively.

@@

\begin{table}
\begin{center}
\begin{tabular}{c c}
\hline \hline
Model, $\tau_{H}$ & $\ V_{max}(km\ s^{-1})$ $0, 50, 100, 200, 300$ \\
\hline \hline
Homogeneous $10^{5}$& $\ 0.2637\  0.2661\  0.3091\  0.3570\  0.3709$ \\
Homogeneous $10^{6}$ & $\ 0.2894\  0.2891\  0.3096\  0.3607\  0.3937$ \\
Homogeneous $10^{7}$ &$\ 0.2279\  0.2298\  0.2318\  0.2547\  0.2813 $\\
Central $10^{5}$ & $\ 0.0966\  0.0966\  0.0965\  0.0965\  0.0965$ \\
Central $10^{6}$ & $\ 0.0663\  0.0665\  0.0665\  0.0665\  0.0665$ \\
Central $10^{7}$ & $\ 0.0149\  0.0158\  0.0158\  0.0158\  0.0158$ \\
\hline
\end{tabular}
\caption{
 Escape fraction values for all of the models. } 
\label{table:escape}
\end{center}
\end{table}

@@

In the right pnael of Figure \ref{fig:efvsv} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{a}$, an approximation that is valid only
in the limit $a\tau_{\rm   H}\gg 1$. The dashed lines in the right panel of Figure
\ref{fig:efvsv} show the results for the different rotational
velocities. First of all we note that the escape fraction does not
increase from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, a fact
that is explained since there is a transition from a opaque to an extremely opaque medium
which affect affects how \ly photons escape from the medium: single flight 
vs. single excursion. {\bf Que quiere decir eso de single flight y
  single excursion, porque explica eso la fraccion de escape?} 
 %we are in a regime where the condition for the analytic expectations
 %($a\tau_{h}\gg 1$) does not hold.

@@

Quantitatively the main results of our study are summarized in the
following items.

@@

\item The line width increases with rotational velocity. The line with
  approximately follows the functional form $W^2 = W_{
    0}^2 + (V_{\rm max}/\lambda)^2$ where $W_{0}$ indicates the line
  width for the static case. $\lambda$ is a constat determined from
  the radiative transfer results it is $\lambda_{\rm c}=3.9\pm$ and
  $\lambda_{\rm h}=1.9\pm$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homgeneously distributed sources. The cases of single
  peaked lines occur when the rotational velocity is larger than the
  line width measured in the static case.

\item For central sources we find that the number of scatterings
does not significantly decrease with rotation, this is due to the 
fact that the vast majority of scatterings events are resonant, 
during which the mean free path is very short and the effect of gas
bulk rotation is not enough to affect \ly photons.
The escape fraction does not depend on the rotational velocity neither
the position of the peaks, but there is an increase 
in the flux on the central parts of the line for high rotational 
velocities.

\item For homogeneous sources we find that the changes in the
line morphology are linked to the reduction in the average number of
scatterings. In the second part of \S \ref{sec:widthpeak} we show
that, as rotational velocity increases, a large fraction of photons
escape with less than $10$ scatterings and on average 
with $40\%$ less scatterings than in the static case. This makes highly improbable 
that a photon's frequency can move away from the line's center. 
In \S \ref{sec:widthpeak} For typical column densities of $N_{HI}=1e^{19}-1e^{20}$, 
we find that rotational velocities of 100-300 \kms broaden the line profile up 
to a factor of $2-3$ in comparison to the static case. Simultaneously, the
flux in the line center also increases with rotational velocity. This
increase induces the transition from a double peak to a single peak
line in the models where the radiation sources are homogeneously
distributed. Also the escape fraction increase by factors of $2-10$ with 
respect to the slab with centrally sources.

\item For non-symmetric sources we find that three peak profiles
could emerge for models in which the rotational velocity is the necessary 
to give rise to a central peak, but not that high to decrease the
amount of photons in the wings. When initial off-center sources are 
selected from the outer parts of the galaxy the wings of the line are stronger.
Also we do not find any considerable asymmetries in the line produced by off-center 
emission.
\end{itemize}

>>>>

We use the full width at half maximum (FWHM) to quantify the line
broadening. Figure \ref{fig:widthvsvelocity} shows how the FWHM increases
with rotational velocity both for the central an homogeneous source
distributions.  We parameterize the dependency of the linewidth as
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.9 \pm 3.17\times 10^{-17}$ and $\lambda_{\rm h}=
1.9\pm 3.07\times 10^{-10}$ respectively.  {\bf EL ERROR ES de $10^{-17}???$}

@@

\begin{table}
\begin{center}
\begin{tabular}{c ccc ccc}
\hline \hline
Sources & $\tau_{H}$ & & & $\ V_{\rm max}$& & \\
& &  &  & (\kms) & & \\ 
& & 0 & 50 & 100 &200 & 300\\ \hline 
Homogeneous & $10^{5}$& 0.263 & 0.266 &  0.309 &  0.357 &  0.370  \\
           & $10^{6}$ & 0.289 &  0.289 &   0.309 &  0.360 &  0.393 \\
         &$10^{7}$ &  0.227 &  0.229 &  0.231 &  0.254 &  0.281 \\
Central & $10^{5}$ &  0.096 & 0.097 & 0.096 &  0.096 & 0.096 \\
  &$10^{6}$ & 0.066 & 0.066 &  0.066 &  0.066 &  0.066 \\
 &$10^{7}$ & 0.015 &  0.016 & 0.016 & 0.016 & 0.015 \\
\hline
\end{tabular}
\caption{
 Escape fraction values for all dusty models. } 
\label{table:escape}
\end{center}
\end{table}

@@

In the right panel of Figure \ref{fig:efvsv} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{a}$, an approximation that is valid only
in the limit $a\tau_{\rm   H}\gg 1$. The dashed lines in the right panel of Figure
\ref{fig:efvsv} show the results for the different rotational
velocities. First of all we note that the escape fraction does not
increase from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, a fact
that is explained since there is a transition from a opaque to an extremely opaque medium
which affect affects how \ly photons escape from the medium: single flight 
vs. single excursion. {\bf Que quiere decir eso de single flight y
  single excursion, porque explica eso la fraccion de escape?} 
 %we are in a regime where the condition for the analytic expectations
 %($a\tau_{h}\gg 1$) does not hold.

@@

Quantitatively the main results of our study are summarized as follows.

@@

\item The line width increases with rotational velocity. The line with
  approximately following the functional for  ${\rm FWHM}^2 = {\rm FWHM}_{
    0}^2 + (V_{\rm max}/\lambda)^2$ where FWHM$_{0}$ indicates the line
  width for the static case. $\lambda$ is a constat determined from
  the radiative transfer results it is $\lambda_{\rm c}=3.9\pm$ and
  $\lambda_{\rm h}=1.9\pm$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homgeneously distributed sources. These cases occur when
  the rotational velocity is larger than hFWHM$_0$.

\item For central sources we find that the number of scatterings
does not significantly decrease with rotation. This is due to the 
fact that the vast majority of scatterings events are resonant, 
during which the mean free path is very short and the effect of gas
bulk rotation is not enough to affect \ly photons. In consequence, 
the escape fraction neither the position of the peaks depende on the
rotational velocity. However, there is a flux increase with rotational
velocity in the central parts of the line.

\item For homogeneous sources we find that the changes in the
line morphology are linked to the reduction in the average number of
scatterings. In the second part of \S \ref{sec:widthpeak} we show
that, as rotational velocity increases, a large fraction of photons
escape with less than $10$ scatterings. The avarage number of
scatterings decreases a $40\%$ compared to the static case. 
The reduction in the number of scattering decreases the probability
for a photon to move away from the line's center. This is also
translated into an increase of the escape fraction by a factor of
$\sim 2$ with respect to the static situation.

\item The spatial anisotropies in the emission are always at a level
  below the $10\%$ deviation from static models.

\item For off-centered emission triple peaked profiles can emerge
  under two conditions: the emission occurs close to th sphere's
  surface and the rotational velocity is half the FWHM in the static
  case.   
\end{itemize}
